Modeling tactical and memory-driven migration in dynamic environments

### 2021-03-25 presentations/memorymigration_v2.Rmd
added final_similarity to runManyRuns
Presentation more or less good to go

**Running the model**

<div class='left' style='float:left;width:48%'>
- R package `deSolve` and `ReacTran` - fluid transport pde solvers.
- Functions all in `memorymigration` GitHub project. 
- SESYNC cluster 
- server-based Rstudio
</div>

<div class='right' style='float:right;width:48%'>

**Assessing the model**

Three metrics:

1. Social cohesiveness
2. Foraging Efficiency
3. Migratoriness
</div>

@@

```{r SC_example, fig.width = 4}
require(memorymigration)
data(resources)
par(mfrow = c(2,1), mar = c(3,3,3,1), mgp = c(2,.25,0), las = 1, tck = 0.01)

@@

$\epsilon = 8; \alpha = 200; \beta_0 = 500; \beta_1 = 0$
![](images/MIexample.png){width=70%}

@@

# Example simulations:

```{r, out.width = "33%", eval = FALSE}
include_graphics(c("images/Sim_withMemory.png","images/Sim_onlyMemory.png", "images/Sim_noMemory.png"))

1. Only Tactics
1. Memory and Sociality
2. Only Memory
3. Only Sociality

@@

<div class='left' style='float:left;width:55%'>

1. Migratoriness results from **concentrated** and **long** resource pulses. 
2. Migratoriness is **increased** by **memory**, but **attenuated** by **sociality**. 
3.  Foraging is easiest (obviously) for **diluted** and **short** resource pulses, as the resource is distributed mainly uniformly.

@@

<div class='right' style='float:right;width:45%'>

![](images/ComparingResourceDynamics_v1.png){width=100%}

@@

>>>>

<div class='left' style='float:left;width:48%'>
## Running the model

- R package `deSolve` and `ReacTran` - fluid transport pde solvers.
- Functions all in `memorymigration` GitHub project. 
- SESYNC cluster 
- server-based Rstudio
</div>

<div class='right' style='float:right;width:48%'>
## Assessing the model

Four metrics:

1. Social cohesiveness
2. Foraging Efficiency
3. Migratoriness
4. Time to equilibrium

@@

```{r SC_example, fig.width = 6}
require(memorymigration)
data(resources)
par(mfrow = c(2,1), mar = c(3,3,3,1), mgp = c(2,.25,0), las = 1, tck = 0.01)

@@

$\epsilon = 8; \alpha = 200; \beta_0 = 500; \beta_1 = 0$

![](images/MIexample.png){width=70%}

@@

# Runs

Lots of them! (8100)

## Model parameters

- $\alpha$ - tactic response: 0, 100, 1000
- $\epsilon$ - diffusion: 10, 30, 100
- $\beta_0$ - sociality: 0, 100, 200, 300, 400, 500
- $\beta_1$ - memory: 0, 100, 200, 300, 400, 500

## Resource pulse parameters

- Duration (2 x sd): 6,12,18,24,30
- Spatial extent (2 x sd): 6,12,18,24,30

# Basic Metrics

<div class='left' style='float:left;width:60%'>
actic response**
2. High overall mobility (high tactic + high diffusion) =  high foraging efficiency
3. Slow diffusion leads to longer time to attain equilibrium

We mainly focus on **low diffusion** and **high tactic**, where most of the (migratory) "action" is.

<div class='right' style='float:right;width:40%'>
![](images/MetricComparisons.png){width=100%}
</div>

# Example simulations

![](images/resource.png){width=20%}![](images/sim0.png){width=70%}

# Example: Memory 0 / Sociality 500

![](images/resource.png){width=20%}![](images/sim2.png){width=70%}

# Example: Memory 500 / Sociality 0

![](images/resource.png){width=20%}![](images/sim1.png){width=70%}

# Example: Memory 200 / Sociality 200

![](images/resource.png){width=20%}![](images/sim3.png){width=70%}

# Example: Memory 500 / Sociality 200 (sometimes things fall apart)

![](images/resource.png){width=20%}![](images/sim4.png){width=70%}

@@

1. Migratoriness results from **concentrated** and **long** resource pulses. 
2. Both migratoriness and sociality **increase migratoriness**. 
3. Foraging efficiency is highest for **diluted** and **short** resource pulses (resource is distributed mainly uniformly).
4. In a "stationary" (equilibrium) scenario, memory and sociality don't contribute much to foraging efficiency (**caveated!**).

@@

![](images/ComparingResourceDynamics.png){width=100%}

@@

## Main take-away

1. This is a convenient, flexible and fast tool for exploring a wide range of memory / tactic / sociality dynamics

## Going forward

"Stress" the system to assess plasticity of a population:

- Make the pulses more isolated (weaken usefulness of tactical strategy).
- Drift the **location** and **timing** of pulse
- Make the pulse more **patchy** and **stochastic**
